---
title: "R Notebook"
output: html_notebook
---
# Introduction

## Background 
The Korean CDC has publically released the country's cases of COVID-19, regularly updated and analyzed most recently on 3/24/20. 

My thoughts: it's way too early, this was before widespread testing maybe? As least I don't understand the disease column, and why the only people labeled "TRUE" have passed. There's no way Korea only tested positive 15 people?? Because from dataset creator: NaN in the 'disease' column means we don't know. That's 98%+ of the data.

This dataset ultimtaely allows us to observe what basic patient demographic features correlate with higher risk of death. This is a classification problem.

```{r}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(ggthemes)
options(scipen = 999)

ori_data <- read.csv('coronavirusdataset/patientinfo.csv', na.strings=c("","NA"))
```

```{r}

## Cleaning
pi <- transform(ori_data, patient_id = as.numeric(patient_id),
                infected_by = as.numeric(infected_by),
                sex = as.factor(sex),
                country = as.factor(country),
                province = as.factor(province),
                city = as.factor(city),
                infection_case = as.factor(infection_case),
                state = as.factor(state),
                symptom_onset_date = as.Date(symptom_onset_date),
                confirmed_date = as.Date(confirmed_date),
                released_date = as.Date(released_date),
                deceased_date = as.Date(deceased_date),
                age = as.factor(age)) %>% 
  select(-global_num, -birth_year, -country) %>% 
  drop_na(state, age, sex, confirmed_date) %>% # remove NAs of key variables for prediction
  unite("outcome_date", released_date:deceased_date, remove = TRUE) # unite released_date and deceased_date

## Clean NA values in outcome_date united column
pi$outcome_date <- gsub("_NA", "", pi$outcome_date)
pi$outcome_date <- gsub("NA_", "", pi$outcome_date)
pi$outcome_date <- gsub("NA", NA, pi$outcome_date)
pi <- pi %>% transform(outcome_date = as.Date(outcome_date))

## Feature engineering 
pi <- pi %>% mutate(days_passed = as.numeric(outcome_date - confirmed_date))

## Repurposing to classifcation problem with binary 'Death' state, 0 = alive, 1 = deceased
pi <- pi %>% mutate(death_outcome = ifelse(state == "deceased", 1, 0)) 
#test_death = pi %>% filter(state == "deceased")

#summary(pi)
#sapply(pi,class)
colSums(is.na(pi))
```

## Descriptive Statistics

```{r}
# Cases per province bar chart
ggplot(pi, aes(x=province, fill = state)) + geom_bar(position="dodge") + ggtitle("Count of Cases per Province") +
  xlab("Province") + ylab("Count")  + theme_light() + theme(axis.text.x = element_text(angle = 90)) + coord_flip()

# Cases of age categories bar chart
ggplot(pi, aes(x=age, fill = state)) + geom_bar(position="stack") + ggtitle("Age Distribution of Cases") +
  xlab("Age") + ylab("Count") + theme_light() 

```

```{r}
## Train test split
index <- sample(1:nrow(pi), nrow(pi)*0.8, replace=F)
train <- pi[index,]
test <- pi[-index,]
```

```{r}
## Model building
library(ranger)
library(rpart)
library(rpart.plot)
library(pROC)
library(car)

# Logit model
logit_model <- glm(death_outcome ~ sex + age + province + confirmed_date, train, family=binomial(logit))
summary(logit_model, results=TRUE)
Anova(logit_model)

# Decision Tree model
DT_model <- rpart(death_outcome ~ sex + age + province + confirmed_date, train, method = "class")
rpart.plot(DT_model, extra=104)

# Random Forest model
train <- train %>% transform(death_outcome = as.factor(death_outcome))
RF_model <- ranger(death_outcome ~ sex + age + province + confirmed_date, train, mtry=2, num.trees = 250, splitrule = "gini",importance = "impurity")

RF.pred <-predict(RF_model, test, type = "response")
RF.test.error <- mean(test$death_outcome!=RF.pred$predictions)
fitRF.roc <-roc(test$state, as.numeric(RF.pred$predictions)) 


```


